Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-18 paper/paper.tex
more edits to paper, this time the mannik section, and updated data Tested everything on quipu.

Huang \emph{et al} ~\cite{huang2003dynamic} has performed a linear
stability analysis on the model which shows an upper limit on a steady
state solution of a $2\mu m$ wavelength.  Cells with dimensions longer
than this are unstable and show oscillatory behavoir in this
direction, while cells that are shorter relax into a motionless steady
state.  We have similarly performed a stability analysis in an
infinite slab with a thickness equal to that of our flattened cells'
($0.25\mu m$) and have found that an equivalent stability limit for
half wavelengths is $2.13\mu m$. As expected, when decreasing the
lengths and widths of our simulated flattened cells so that the
longest distance across the cell is less than this length, the cells
stop exhibiting any oscillalitory behavior.  The deterministic
simulations relax into a motionless state and the stochastic
simulations exhibit multiple random locations of small maximization.

@@

As describe in Section~\ref{sec:model-method-shapes}, we will focus on
simulation of four flattened cell shapes: two shapes copied from
Mannik's experiment~\cite{mannik2012robustness} (shape~\emph{A} and
shape~\emph{B}), and for comparison two symmetrical stadium shapes
that have the identical aspect ratio, thickness and volume (stadium
\emph{A} and stadium \emph{B}).  These stadium shapes enable us to
distinguish between the effect of cellular irregularity and that of
overall size and aspect ratio.  We note that shape \emph{B} is more
irregular than shape \emph{A}, and in particular features a prominant
region of concavity on its left-hand side.  For each of these shapes
(in addition to the wild type pill shape discussed previously) we have
simulated in excess of \fixme{???} seconds of evolution of the MinD
system using both the deterministic model of Huang~\fixme{cite} and
the stochastic model of Huang~\fixme{cite}.

From each of these flattened simulations, we have chosen a typical 350
second segment to compare with the results of Mannik exemplified in
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell (over the same time
interval)~\cite{mannik2012robustness}.  This comparison is shown in
Fig.~\ref{randst-plot-ave}.  In addition to arrows between successive
maxima in space and time, we plot as a colored background the density
of MinD protiens averaged over the same time period.  We note that we
have manually verified that our (computer-generated) arrow plots also
reflect a human interpretation of a movie of the same data.  Finally,
for comparison we present the same plot for a wild-type cell, with a
time period of \fixme{???}  seconds to account for its short period of
oscillation.  In every case, including the wild type pill shape, we
see irregularity in the location of the maxima when using the
stochastic approach.  The deterministic method shows uniformly bipolar
oscillation, with the exception of shape \emph{B}, which develops weak
intermediate maxima on the right-hand edge of the cell.

@@

The predictions of the stochastic as seen in
Fig.~\ref{randst-plot-ave} are roughly similar to the experimental
results in terms of irregularity of the locations of maxima, although
there are locations on the edges of the cells where experiment shows
maxima occuring that we never see in our simulations, suggesting that
the model does not precisely reflect the experiment.  We also note
that the stadium shapes in Fig.~\ref{randst-plot-ave} appear
qualitatively similar to the irregular shapes in terms of the
locations of their maxima.  From these results, we conclude that the
deterministic method is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

>>>>

Huang \emph{et al}~\cite{huang2003dynamic} has performed a linear
stability analysis on the model which shows an upper limit on a steady
state solution of a $2\mu m$ wavelength.  Cells with dimensions longer
than this are unstable and show oscillatory behavoir in this
direction, while cells that are shorter relax into a motionless steady
state.  We have similarly performed a stability analysis in an
infinite slab with a thickness equal to that of our flattened cells'
($0.25\mu m$) and have found that an equivalent stability limit for
half wavelengths is $2.13\mu m$. As expected, when decreasing the
lengths and widths of our simulated flattened cells so that the
longest distance across the cell is less than this length, the cells
stop exhibiting any oscillalitory behavior.  The deterministic
simulations relax into a motionless state and the stochastic
simulations exhibit multiple random locations of small maximization.

@@

As describe in Section~\ref{sec:model-method-shapes}, we will focus on
simulation of four illustrative flattened cell shapes: two shapes
created to replicate those published by
Mannik~\cite{mannik2012robustness} (shape~\emph{A} and
shape~\emph{B}), and for comparison two symmetrical stadium shapes
that have the identical aspect ratio, thickness and volume (stadium
\emph{A} and stadium \emph{B}).  These can be seen in
Figure~\ref{randst-plot-ave}.  These stadium shapes enable us to
distinguish between the effect of cellular shape irregularity and
assymetry, and that of overall size and aspect ratio.  We note that
shape \emph{B} is more irregular than shape \emph{A}, and in
particular features a prominant region of concavity on its left-hand
side.  For each of these shapes (in addition to the wild type pill
shape discussed previously) we have simulated in excess of \fixme{???}
seconds of evolution of the MinD system using the deterministic method
of Huang's model~\cite{huang2003dynamic}, as well as the stochastic
method.

From each of these flattened simulations, we have chosen a typical 350
second segment to compare with the results of Mannik exemplified in
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell (over the same time
interval)~\cite{mannik2012robustness}.  This comparison is shown in
Fig.~\ref{randst-plot-ave}.  In addition to arrows between successive
maxima in space and time, we plot as a colored background the density
of MinD protiens averaged over the same time period.  We note that we
have manually verified that our (computer-generated) arrow plots also
reflect a human interpretation of a movie of the same data.  Finally,
for comparison we present the same plot for a wild-type cell, with a
time period of \fixme{???}  seconds to account for its short period of
oscillation.

In every case, including the wild type pill shape, we see irregularity
in the location of the maxima when using the stochastic approach.  The
deterministic method shows uniformly bipolar oscillation, with the
exception of shape \emph{B}, which develops weak intermediate maxima
on the right-hand edge of the cell.  From these results, we conclude
that the deterministic method is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

@@

The predictions of the stochastic as seen in
Fig.~\ref{randst-plot-ave} are roughly similar to the experimental
results in terms of irregularity of the locations of maxima, although
there are locations on the edges of the cells where experiment shows
maxima occuring that we never see in our simulations, suggesting that
the model does not precisely reflect the experiment.  We also note
that the stadium shapes in Fig.~\ref{randst-plot-ave} appear
qualitatively similar to the irregular shapes in terms of the
locations of their maxima.
